Implementation of a lattice Boltzmann method for immiscible two-phase flow simulations using the level set method

### 2015-11-15 Paper/paper.tex
Lid driven cavity
mit tropfen

\section{Introduction}
%Something about the two-phase problem and some other algorithms that solve it. We could do this according to the presentation-slides.
In engineering the immiscible multiphase flow problem often needs to be considered. For example, bubble dynamics are decisive for the design of chemical reactors. Also, fingering in oil recovery is a common multiphase problem.

@@

Figure \ref{laplace} shows the simulation results we obtained for a $40 \times 40$ cells grid over a $[0,1]\times[0,1]$-domain with periodic boundaries. At the center, a sphere of radius $0.25$ was placed with $\varrho_2 = 1.1$ while the exterior contained a fluid with $\varrho_1 = 1$. It is visible, how the approximated values attain the analytical solution in a oscillatory motion decaying amplitude, until both values agree. The algorithm needed 1000 LBM-iterations, where every ten iterations, a call to the level-set routine was executed. Both fluids had an identical viscosity of $\nu = 1/6$.
\begin{figure}[t!]
	\flushright
	\hfill\includegraphics[trim = 0mm 7.75cm 0mm 8.06cm, clip, width=.8\textwidth, natwidth=595,natheight=842]{laplace.pdf}\hspace*{\fill}
	\caption{The pressure drop of a spherical bubble over LBM iterations}
	\label{laplace}
\end{figure}
We noted that the velocity close to the interface contained erroneous components, which lead to a slight distortion of the spherical shape. We explain those fluxes as inaccuracies stemming from the used level set subroutines, according to the analysis shown in \cite{Thoemmes2}.
Additionally, we want to state, that we were not able to attain stable results for surface tension coefficients $\sigma > \mathrm{O(10^{-3})}$. Large densitiy ratios were limited to big viscosity regimes, or refined spatial resolution in our tests.

\section{Conclusion}
We implemented a hybrid lattice Boltzmann method that uses the level set method to model the movement of the interface. The implementation was done in Matlab for a two dimensional case. We adapted the formulas given in \cite{Thoemmes} to fit in our environment. Finally, we verified our implementation with different test cases.
Possible applications for our algorithm are shown in figures TODO \& TODO, which show the transient movement of an oil drop in a lid driven cavity. %Fig TODO shows two rising and coalescing bubbles in a fluid, produced with our code.

We particularly want to thank Ian Mitchell for providing the toolbox for level set methods.

@@

>>>>

\section{Introduction}
In engineering the immiscible multiphase flow problem often needs to be considered. For example, bubble dynamics are decisive for the design of chemical reactors \cite{trygg}. Also, fingering in oil recovery is a common multiphase problem.

@@

Figure \ref{laplace} shows the simulation results we obtained for a $40 \times 40$ cells grid over a $[0,1]\times[0,1]$-domain with periodic boundaries. At the center, a sphere of radius $0.25$ was placed with $\varrho_2 = 1.1$ while the exterior contained a fluid with $\varrho_1 = 1$. It is visible, how the approximated pressure attains the analytical solution in a oscillatory motion with decaying amplitude, until both values agree. The algorithm needed 1000 LBM-iterations, where every ten iterations, a call to the level-set routine was executed. Both fluids had an identical viscosity of $\nu = 1/6$.
\begin{figure}[t!]
	\flushright
	\hfill\includegraphics[trim = 0mm 7.75cm 0mm 8.06cm, clip, width=.8\textwidth, natwidth=595,natheight=842]{laplace.pdf}\hspace*{\fill}
	\caption{The pressure drop of a spherical bubble over LBM iterations}
	\label{laplace}
\end{figure}
We noted that the velocity close to the interface contained erroneous components, which lead to a slight distortion of the spherical shape. We explain those fluxes as inaccuracies stemming from the used level set subroutines, according to the analysis shown in \cite{Thoemmes2}.
Additionally, we want to state, that we were not able to attain stable results for surface tension coefficients $\sigma > \mathrm{O(10^{-3})}$. Large densitiy ratios were limited to big viscosity regimes, or refined spatial resolution in our tests.

\section{Conclusion}
We implemented a hybrid lattice Boltzmann method that uses the level set method to model the movement of the interface. The implementation was done in Matlab for a two dimensional case. We adapted the formulas given in \cite{Thoemmes} to fit in our environment. Finally, we verified our implementation with different test cases.

A more vivid possible application of the algorithm we presented is shown in figure TODO, which shows the transient movement of an oil drop in a lid driven cavity. %Fig TODO shows two rising and coalescing bubbles in a fluid, produced with our code.

The authors want to thank Ian Mitchell for providing the toolbox for level set methods.

@@

\bibitem{trygg} {\sc G. TRYGGVASON, R. SCARDOVELLI, S. ZALESKI}, Direct Numerical Simulations of Gas-Liquid Multiphase Flows, Cambridge University Press, ISBN 9780521782401 (2011)
